---
title: "R coding algorithms for migrating distance quantification"
output: html_document
---

#### Transretinal migration of astrocytes and brain/spinal cord-like cells arising from subretinally transplanted human retinal organoids

Ying V. Liu†, Clayton P. Santiago†, Akin Sogunro, Gregory K. Konar, Ming-Wen Hu, Minda M. McNally, Yuchen Lu, Zhuolin Li, Dzhalal Agakishiev, Sarah Hadyniak, Katarzyna Hussey, Tyler J. Creamer, Linda D. Orzolek, Derek Teng, Jiang Qian, Zheng Jiang, Robert J. Johnston, Jr.#, Seth Blackshaw#, and Mandeep S. Singh1#


Load required packages
```{r setup, eval=FALSE}
library(data.table)
library(plyr)
```

Extracting data exported from ImageJ
```{r, eval=FALSE}
# Extract the X,Y coordinate of migrated cells, 
# and generate subsequent distance data presented in Table 1
common_path = "~/Desktop/cell invasion/Processed_Image"

files_to_read = list.files(
  path = common_path,        # directory to search within
  pattern = ".*(Rd1-NS).*csv$", # regex pattern
  recursive = TRUE,          # search subdirectories
  full.names = TRUE          # return the full path
)
```

Set up dataframe
```{r, eval=FALSE}
# Hypothetical Researcher has 37 retina slides to quantify 
# and wants to localize 5 cell types per retina slides. So that's 
# 37 rows and 6 columns (including cell id) in the data list

data_lst = lapply(files_to_read, read.csv)  # read all the matching files
celltype_summary = data.frame(matrix(ncol = 6, nrow = 37))
colnames (celltype_summary) <- c("File_name", "Type_1", "Type_2", "Type_3", "Type_4", "Type_5")
```

Calculating distance
```{r, eval=FALSE}
# Calculating the distance from starting point for each cell 
# categorized by cell types 

for (i in 1:length(data_lst)){
  data_lst[[i]]$Address <- rep(files_to_read[i],nrow(data_lst[[i]]))
  File_Name <-  files_to_read[i]
  Type_1 <- sum(data_lst[[i]]$Type == 1)
  Type_2 <- sum(data_lst[[i]]$Type == 2)
  Type_3 <- sum(data_lst[[i]]$Type == 3)
  Type_4 <- sum(data_lst[[i]]$Type == 4)
  Type_5 <- sum(data_lst[[i]]$Type == 5)
  celltype_summary[i,1:6] = c(File_Name, Type_1, Type_2, Type_3, Type_4, Type_5)
  print(c(File_Name, Type_1, Type_2, Type_3, Type_4, Type_5))
  for (j in 1:nrow(data_lst[[i]])){
    data_lst[[i]][j,"Z.µm."]=sqrt(( data_lst[[i]][j,"X.µm."]- data_lst[[i]][1,"X.µm."])^2 +
                                    ( data_lst[[i]][j,"Y.µm."]- data_lst[[i]][1,"Y.µm."])^2) #distance calculation
  }
}
```

Exporting analyzed invasion distance summary
```{r, eval=FALSE}
dat1<-ldply(data_lst)

write.table(as.data.frame(dat1),file="Detailed_Result.csv", quote=F,sep=",",row.names=F)
write.table(as.data.frame(celltype_summary),file="Celltype_Summary.csv", quote=F,sep=",",row.names=F)
```
